[Re] Speed/accuracy trade-off between the habitual and the goal-directed processes

### 2016-02-09 article/viejo_girard_khamassi.tex
Update after first round of reviewing

\begin{longtable}[c]{@{}lll@{}}
\toprule
Name & Description & Value\tabularnewline
\midrule
\endhead
\(\sigma\) & Updating rate of the average reward & 0.02\tabularnewline
\(\eta\) & Variance of evolution noise & 0.0001\tabularnewline
\(P_n\) & Variance of observation noise & 0.05\tabularnewline
\(\beta\) & Rate of exploration & 1.0\tabularnewline
\(\rho\) & Update rate of the reward function & 0.1\tabularnewline
\(\gamma\) & Discount factor & 0.95\tabularnewline
\(\tau\) & Time step of graph exploration & 0.08\tabularnewline
depth & Depth of search in graph exploration & 3\tabularnewline
\(\phi\) & Update rate of the transition function & 0.5\tabularnewline
init cov & Initialisaton of covariance matrice & 1.0\tabularnewline
\(\kappa\) & Unscentered transform parameters & 0.1\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{quote}
\begin{quote}
Initialization
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\[Q(s, a)^{Goal-Directed} = \{ 0, \ldots \}\]
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\[Q(s, a)^{Habitual} = \{ 0, \dots \}\]
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\[
\Sigma = \left(
    \begin{array}{*4{c}}
    cov \times \eta &  0 &  \ldots & 0 \\
     0 &  cov \times \eta &  \ldots & \vdots \\
    \vdots &  \ldots &  \ddots & 0 \\
     0 &  \ldots &  0 & cov \times \eta \\     
  \end{array}\right)
\]
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(R(S1, EM) = 1\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\bar{R} = 0\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\hat{R}(s,a) = \{0,\ldots\}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
Main Loop
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{FOR}\ i = 1:T\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(s_t = S_0\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{IF}\ i = T_{devaluation}\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(R(S1, EM) = 0\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\hat{R}(S1, EM) = -1\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{WHILE}\ s_t \neq S1 \bigwedge a_t \neq EM\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(a_t = \textbf{Selection}(s_t)\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(r_t = R(s_t,a_t)\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(s_{t+1} = transition(s_t, a_t)\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{Update}(s_t,a_t, s_{t+1}, r_t)\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
Selection
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\{a_1,\ldots,ai,\ldots\} \leftarrow sort(Q(s_t,a_i))\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(VPI(s_t, a_1) = (Q(s_t,a_2)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_1)^{H}<Q(s_t,a_2)^{H}) + \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_2)^H - Q(s_t,a_1)^H)^2}{2\sigma(s_t,a_t)^2}}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(VPI(s_t, a_i) = (Q(s_t,a_i)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_i)^{H}>Q(s_t,a_1)^{H}) + \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_1)^H - Q(s_t,a_i)^H)^2}{2\sigma(s_t,a_t)^2}}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{FOR}\ i \in \{a_1, a_2,\ldots, a_i, \ldots\}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{IF}\ VPI(s_t, a_i) \geq \tau \bar{R}\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(Q(s_t,a_i) = \hat{R}(s_t,a_i) + \gamma \sum\limits_{s'}p_{T}(\{s,a\}\rightarrow s') \max\limits_{b \in A} Q(s',b)^{Goal-directed}\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{ELSE}\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(Q(s_t,a_i) = Q(s_t,a_i)^{Habitual}\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(a_t \leftarrow \textit{SoftMax}(Q(s_t,a), \beta)\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
Update
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\bar{R} = (1-\sigma) \bar{R} + \sigma r_t\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\hat{R}(s_t,a_t) =(1 - \rho) \hat{R} + \rho r_t\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(p_{T}(s_t, a_t, s_{t+1}) = (1 - \phi) p_{T}(s_t, a_t, s_{t+1}) + \phi\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
Specific to Kalman Q-Learning
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\Theta = \{ \theta_j, 0 \geq j \geq 2|S.A|\}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\check{W} = \{ w_j, 0 \geq j \geq 2|S.A| \}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\check{R} = \{ \check{r}_j = \theta_j(s_t,a_t) - \gamma \max\limits_{b \in A} \theta_j(s_{t+1},b),\ 0 \geq j \geq 2|S.A|\}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(r_{predicted} = \sum\limits_{j=0}^{2|S.A|} w_j \check{r}_j\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(P_{\theta_j \check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\theta_j - Q^{Habitual}_t)(\check{r}_j - r_{predicted})\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(P_{\check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\check{r}_j - r_{predicted})^2 + P_n\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(K_t = P_{\theta_j \check{r}_j} P_{\check{r}_j}^{-1}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\delta_t = r_t - r_{predicted}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(Q_{t+1}^{Habitual} = Q_{t}^{H} + K_t \delta_t\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(P_{t+1}^{H} = P_{t}^{H} - K_t P_{\Sigma_t} K_t^T\)
\end{quote}
\end{quote}
\end{quote}

>>>>

\begin{longtable}[c]{@{}lll@{}}
\toprule
Name & Description & Value\tabularnewline
\midrule
\endhead
\(\sigma\) & Updating rate of the average reward & 0.02\tabularnewline
\(\eta\) & Variance of evolution noise & 0.0001\tabularnewline
\(P_n\) & Variance of observation noise & 0.05\tabularnewline
\(\beta\) & Rate of exploration & 1.0\tabularnewline
\(\rho\) & Update rate of the reward function & 0.1\tabularnewline
\(\gamma\) & Discount factor & 0.95\tabularnewline
\(\tau\) & Time step of graph exploration & 0.08\tabularnewline
depth & Depth of search in graph exploration & 3\tabularnewline
\(\phi\) & Update rate of the transition function & 0.5\tabularnewline
init cov & Initialisation of covariance matrice & 1.0\tabularnewline
\(\kappa\) & Unscentered transform parameters & 0.1\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{quote}
\textbf{Initialization}
\end{quote}

\begin{quote}
\begin{quote}
\(Q(s, a)^{Goal-Directed} = \{ 0, \ldots \}\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(Q(s, a)^{Habitual} = \{ 0, \dots \}\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\#\) Covariance matrix
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\Sigma = \left(  \begin{array}{*4{c}}  cov \times \eta & 0 & \ldots & 0 \\  0 & cov \times \eta & \ldots & \vdots \\  \vdots & \ldots & \ddots & 0 \\  0 & \ldots & 0 & cov \times \eta \\  \end{array}\right)\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(R(S1, EM) = 1\) \(\#\) Reward value
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\bar{R} = 0\) \(\#\) Reward rate
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\hat{R}(s,a) = \{0,\ldots\}\) \(\#\) Reward function
\end{quote}
\end{quote}

\begin{quote}
\textbf{Main Loop}
\end{quote}

\begin{quote}
\begin{quote}
\(\textbf{FOR}\ i = 1:T\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(s_t = S_0\) \(\#\) Initial state
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{IF}\ i = T_{devaluation}\) \(\#\) Moderate / Extensive
training
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(R(S1, EM) = 0\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\hat{R}(S1, EM) = -1\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{WHILE}\ s_t \neq S1 \bigwedge a_t \neq EM\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(a_t = \textbf{Selection}(s_t)\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(r_t = R(s_t,a_t)\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(s_{t+1} = transition(s_t, a_t)\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{Update}(s_t,a_t, s_{t+1}, r_t)\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\textbf{Selection}
\end{quote}

\begin{quote}
\begin{quote}
\(\#\) Sort the Q-values in descending order
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\{a_1,\ldots,ai,\ldots\} \leftarrow sort(Q(s_t,a_i))\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\#\) VPI : Value of Precise Information
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(VPI(s_t, a_1) = (Q(s_t,a_2)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_1)^{H}<Q(s_t,a_2)^{H})+ \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_2)^H - Q(s_t,a_1)^H)^2}{2\sigma(s_t,a_t)^2}}\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(VPI(s_t, a_i) = (Q(s_t,a_i)^{H}-Q(s_t,a_1)^{H})P(Q(s_t,a_i)^{H}>Q(s_t,a_1)^{H}) + \frac{\sigma(s_t,a_t)}{\sqrt{2\pi}} e^{-\frac{(Q(s_t,a_1)^H - Q(s_t,a_i)^H)^2}{2\sigma(s_t,a_t)^2}}\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\textbf{FOR}\ i \in \{a_1, a_2,\ldots, a_i, \ldots\}\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{IF}\ VPI(s_t, a_i) \geq \tau \bar{R}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\#\) Q-Value from Goal-directed system is evaluated
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(Q(s_t,a_i) = \hat{R}(s_t,a_i) + \gamma \sum\limits_{s'}p_{T}(\{s,a\}\rightarrow s') \max\limits_{b \in A} Q(s',b)^{Goal-directed}\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\(\textbf{ELSE}\)
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(\#\) Q-Value from Habitual system is retrieved
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\(Q(s_t,a_i) = Q(s_t,a_i)^{Habitual}\)
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(a_t \leftarrow \textit{SoftMax}(Q(s_t,a), \beta)\)
\end{quote}
\end{quote}

\begin{quote}
\textbf{Update}
\end{quote}

\begin{quote}
\begin{quote}
\(\bar{R} = (1-\sigma) \bar{R} + \sigma r_t\) \(\#\) Reward Rate
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\hat{R}(s_t,a_t) =(1 - \rho) \hat{R} + \rho r_t\) \(\#\) Reward
function
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(p_{T}(s_t, a_t, s_{t+1}) = (1 - \phi) p_{T}(s_t, a_t, s_{t+1}) + \phi\)
\(\#\) Probability of transition
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
Specific to Kalman Q-Learning
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\#\) Sigma-points sampling
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\Theta = \{ \theta_j, 0 \geq j \geq 2|S.A|\}\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\check{W} = \{ w_j, 0 \geq j \geq 2|S.A| \}\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\check{R} = \{ \check{r}_j = \theta_j(s_t,a_t) - \gamma \max\limits_{b \in A} \theta_j(s_{t+1},b),\ 0 \geq j \geq 2|S.A|\}\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(r_{predicted} = \sum\limits_{j=0}^{2|S.A|} w_j \check{r}_j\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\#\) Covariance computation
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(P_{\theta_j \check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\theta_j - Q^{Habitual}_t)(\check{r}_j - r_{predicted})\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(P_{\check{r}_j} = \sum\limits_{j=0}^{2|S.A|} w_j (\check{r}_j - r_{predicted})^2 + P_n\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(K_t = P_{\theta_j \check{r}_j} P_{\check{r}_j}^{-1}\) \(\#\) Kalman
gain
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(\delta_t = r_t - r_{predicted}\) \(\#\) Reward-prediction error
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(Q_{t+1}^{Habitual} = Q_{t}^{H} + K_t \delta_t\)
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
\(P_{t+1}^{H} = P_{t}^{H} - K_t P_{\Sigma_t} K_t^T\)
\end{quote}
\end{quote}
